---
title: "Psychometric Evaluation and Validation of Bangla Rotter Internal-External Scale with Classical Test Theory and Item Response theory"
author: "Mushfiqul Anwar Siraji"
date: "`r Sys.Date()`"
output: 
  bookdown::html_document2: 
    toc: yes
    toc_float:
      toc_collapsed: yes
    number_sections: yes
    theme: lumen
    highlight: tango
bibliography: Rotter-References.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)
set.seed(123)
```

# Preparing the R environment

```{r install-pack, eval=FALSE}
# Run this to install the required packages.
install.packages("pacman")
pacman::p_load(devtools,tidyverse, psych, mirt, lavaan, VIM, semPower, labelled, MOTE, semTools, knitr, kableExtra)
devtools::install_github("crsh/papaja", force = T)
devtools::install_github('SachaEpskamp/semPlot',  dependencies = T)
remotes::install_github("ddsjoberg/gtsummary")

```

```{r load-libs, echo=TRUE, results='hide'}
library(tidyverse)
library(psych)
library(mirt)
library(lavaan)
library(gtsummary)
library(VIM)
library(tabledown)
library(semPower)
library(labelled)
library(MOTE)
library(semTools)
library(papaja)
library(knitr)
library(kableExtra)
papaja::r_refs("Rotter-References.bib")
```

# Sample size estimation
- For EFA generally a sample size of 250-300 is required [@RN1249; @RN1250].
- For CFA we should follow n:q rule where at least 5/10/20 participants are required per item [@R-msm; @RN1269; @RN1270; @RN1271].
- For IRT it should be around 500. But should use a monte carlo simulation for sample size estimation.

```{r samplesize-estimation}
lavmodel <-  "LOC =~ item18 + item25 +item11 + item15 + item16 + item10 +item9 + 
item23 + item13 + item5 + item28 + item4" 
df <- semPower::semPower.getDf(lavmodel)

ap <- semPower::semPower.aPriori(effect = .05, effect.measure = 'RMSEA', alpha = .05, 
                       power = .80, df = df) 
summary(ap)

```

# Data analysis 

## Load and subset the data

```{r data}

#Load the data
data_0 <- readRDS("WRotter.rds")
data <- labelled::remove_labels(data_0)
##EFA data
Rotter.EFA <- data[1:300,11:33]
##EFA Descriptivs
efa.descriptives <- data[1:300,c(3,4,5,6,7,8)]
##CFA Data
Rotter.CFA <- data[301:478, 11:33]
##CFA Descriptives
cfa.descriptives <- data[301:478,c(3,4,5,6,7,8)]
##IRT data
WRotter.IRT <- data[,11:33]
##IRT Descriptives
irt.descriptives <- data[,c(3,4,5,6,7,8)]

## Data for descriptives (Publication)
Descriptives.pub <- data %>% 
  dplyr::select("Gender", "Age", "Education Years", "Social Stance","Marital Status","sample")
```

## Cheack missing values

```{r missing-values}
#Check Missing Values
missing.EFA <- VIM::aggr(Rotter.EFA, plot =T)
missing.CFA<- VIM::aggr(Rotter.CFA, plot =T)
missing.IRT <- VIM::aggr(WRotter.IRT, plot =T)

```
Our data has no missing values. 

## Descriptive statistics

There are several ways of getting the descriptive summary. At first we will do it using **psych**[@R-psych] package to get item-summary.

```{r des-psych}

Descriptives <- psych::describe(Rotter.EFA)
Descriptives %>% 
  knitr::kable(caption = "Descriptives", col.names = NA, digits = 2) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F) %>% 
    scroll_box(width = "100%", height = "450px")
```

**Skewness >2.0** and **Kurtosis > 7.0** indicate severe uni-variate non normality. 


Now we will try **gtsummary** [@R-gtsummary] package to get descriptive statistics of the sample.

```{r gtsum, eval=FALSE, include=FALSE}
desc_sample <-  gtsummary::tbl_summary(Descriptives.pub,
                                      by = sample,
                                      statistic = list(
                                        all_continuous() ~ "{mean} ({sd})",
                                        all_categorical() ~ "{n} ({p}%)"),
                                      type = list(`Education Years` ~ 'continuous'),
                                      digits = all_continuous() ~ 2,
                                      missing = "no") %>%
  # modify_caption("**Descriptive Statistics of the Sample:1**") %>% 
  add_overall() %>% add_p()


desc_gender <-  gtsummary::tbl_summary(Descriptives.pub,
                                      by = Gender,
                                      statistic = list(
                                        all_continuous() ~ "{mean} ({sd})",
                                        all_categorical() ~ "{n} ({p}%)"),
                                      type = list(`Education Years` ~ 'continuous'),
                                      digits = all_continuous() ~ 2,
                                      missing = "no") %>%add_overall()%>% add_p()
  # modify_caption("**Descriptive Statistics of the Sample:2**") %>%

```


## Assumptions
### Univariate normality
Now we will try to look into item summary using **tabledown** package. Here we will check the univariate normality

```{r sim-tabledown}
des.tab <- tabledown::des.tab(Rotter.EFA)

papaja::apa_table(des.tab, caption= "A beautiful table describing the item summary ")
```

### Multivariate normality

```{r mn}
mardia <- mardia(Rotter.EFA, na.rm = T, plot =T)

```

## Corrleaton matrix
```{r correlation}
correlations <- psych::tetrachoric(Rotter.EFA,smooth=TRUE)
upper <- correlations$rho[upper.tri(correlations$rho, diag = F)]

range.cor <- range(upper)

##Quality of the matrix
bartlet <- psych::cortest.bartlett(Rotter.EFA, n =300)

```

```{r correlation-plot}
p.mat <- correlation::cor_to_p(correlations$rho, 300, method = "tetrachoric")

corplot <- ggcorrplot::ggcorrplot(correlations$rho, p.mat=p.mat$p, insig = "pch", hc.order = FALSE, outline.col = "white", type = "lower", legend.title = "Correlation",sig.level = 0.05, pch.cex=1.8,
                                  # insig = "blank",
                                  ggtheme = ggplot2::theme_minimal(),tl.srt = 90,colors = c("#E64B35FF", "white", "#3C5488FF") )+
  theme( panel.grid.major = element_blank(),
         axis.text.x = element_text(size = 8),
         axis.text.y = element_text(size = 8),
         panel.border = element_rect(color = "black",
                                     fill = NA,
                                     size = 1))

corplot
```

## Item analysis: classical test theory
Before starting the EFA, executing an item analysis will give us idea about the quality of the items.  

```{r Cronbach, results='asis', width = 60}
Cron.bach <- psych::alpha(Rotter.EFA, check.keys = T) 
Cron.bach$total %>% 
  knitr::kable(caption = "Alpha", digits = 2) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)
Cron.bach$alpha.drop %>% 
  knitr::kable(caption = "Alpha value if item is dropped", digits = 2) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)

Cron.bach$item.stats %>% 
  knitr::kable(caption = "Corrected Item total Correlation", digits = 2) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F)
  
```

## Factor analysis

## Sampling adequacy: KMO

```{r KMO}
KMO <- KMO(Rotter.EFA) 
kmo.value <- KMO$MSA
```

Minimum cut off is **.50** [@RN1221]. Our Overall value is `r MOTE::apa(kmo.value,2,T)`.

```{r KMO-itemwise}
KMO$MSAi%>% 
knitr::kable(caption = "Item-Wise Value", digits = 2) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = F) %>% 
  scroll_box(width = "200px", height = "450px")

min.kmo <- min(KMO$MSAi)
max.kmo <- max(KMO$MSAi)
```

Individual items KMO value ranges between `r apa(min.kmo,2,T)`- `r apa(max.kmo,2,T)`.

## Identification of factors

```{r scree}
scree(correlations$rho,factors=TRUE,pc=TRUE,main="Scree Plot",
               hline=NULL,add=F)
```

```{r hull}
EFA.MRFA::hullEFA(Rotter.EFA,extr = "ULS", index_hull = "CAF", display = TRUE, 
                  graph = T,
                  details = TRUE)
```

## Factor analysis
```{r fa, include =F}
fa.2F.1 <- fa(r=correlations$rho, nfactors = 2, fm= "wls", rotate = "varimax",
              residuals = TRUE, SMC = TRUE, n.obs =300)
print(fa.2F.1, cut = .3, digits = 3, sort = T)

#Refitting the model
reduced.model.2F.1 <- dplyr::select(Rotter.EFA, -item22, -item6,-item29,-item25, -item9)
correlations.redu.2F.1 <-  tetrachoric(reduced.model.2F.1)
fa.2F.2 <- fa(r=correlations.redu.2F.1$rho, nfactors = 2, fm= "wls", rotate = "varimax",
              residuals = TRUE, SMC = TRUE, n.obs =300)
print(fa.2F.2, cut = .3, digits = 3, sort = TRUE)

Factor_analysis_tab <- FAc <-tabledown::fac.tab( fa.2F.2, .3, complexity = F)

```

```{r fa-tab}
Factor_analysis_tab <- FAc <-tabledown::fac.tab( fa.2F.2, .3, complexity = F)

papaja::apa_table(Factor_analysis_tab, caption= "A beautiful table describing the fator analysis output ")
```

## Confirmatory factor analysis

```{r cfa}
Rotter.one <- "LOC =~ item18 + item25 +item11 + item15 + item16 + item10 +item9 + 
item23 + item13 + item5 + item28 + item4"

Rotter.one.fit <- cfa(Rotter.one,  Rotter.CFA,  mimic = "Mplus",
                      ordered = names(Rotter.CFA),estimator = "WLSMV")

One.factor.sum <- summary(Rotter.one.fit, fit.measures =TRUE,standardized = TRUE,
                          rsq =TRUE)

One.factor.Cfa <- fitmeasures (Rotter.one.fit,c("gfi", "agfi", "nfi","rfi", 
                                                "cfi.scaled","tli.scaled",
                                                "rmsea", 
                                                "srmr","aic"))



One.factor.reliability <- semTools::reliability(Rotter.one.fit)
# modfit.Cor.one <- modindices(Rotter.one.fit, sort. = TRUE) 
# modfit.Cor.one[modfit.Cor.one$mi>3.84,]


Rotter.one.2 <- "LOC =~ item18 + item25 +item11 + item15 + item16 + item10 +item9 + item13 + item5 + item28 + item4"

Rotter.one.fit.2 <- cfa(Rotter.one.2,  Rotter.CFA,  mimic = "Mplus",
                        ordered = names(Rotter.CFA),estimator = "WLSMV")

One.factor.Cfa.mod.sum <- summary(Rotter.one.fit.2, fit.measures =TRUE,standardized = TRUE, rsq =TRUE)

One.factor.Cfa.mod.fit <- fitmeasures (Rotter.one.fit.2,c("gfi", "agfi", "nfi","rfi", 
                                                          "cfi.scaled","tli.scaled",
                                                          "rmsea", 
                                                          "srmr","aic"))


one.factor.cfaindices <- as.data.frame(One.factor.Cfa)
fitindices <- rownames_to_column(one.factor.cfaindices)
one.factor.cfaindices.mod <- as.data.frame(One.factor.Cfa.mod.fit)
cfatab1 <- cbind(fitindices,one.factor.cfaindices.mod   )
cfatab2 <- as.data.frame(t(cfatab1))
cfatab3 <- cfatab2[-1,-9]

One.factor.reliability.mod <- semTools::reliability(Rotter.one.fit.2)

# cfatab4 <- cbind(cfatab3,One.factor.reliability[5],One.factor.reliability.mod[5])
rownames(cfatab3) <- c("Model1: 12 items", "Model2: 11 items")
cfatab3$Reliability[1] <- One.factor.reliability[5]
cfatab3$Reliability[2] <- One.factor.reliability.mod[5]
```

```{r cfa-tab}
apa_table(cfatab3, caption= "CFA Fit Indices")
```


```{r Cfa-plot}
CFA_plot1 <- semPlot::semPaths (Rotter.one.fit.2 , 
                                  what= "std",
                                  intercepts = F,
                                  style ="OpenMx",
                                  theme = "colorblind",
                                  nCharNodes = 0,
                                  reorder =T,
                                  rotation =2,
                                  layout ="tree",
                                  cardinal = T,
                                  curvePivot =T,
                                  sizeMan =6,#items length
                                  sizeMan2 = 2,#items
                                  sizeLat = 12,#factors
                                  thresholds = FALSE,
                                  equalizeManifests =F,
                                  fade = FALSE)

```

## Item analysis: item response theory

```{r IRT, include =F}
#IRT
Rotter.one.IRT <- WRotter.IRT %>% 
  dplyr::select(item18, item25, item11, item15, item16, item10, item9, item13,  
                item5,  item28, item4)


unidimensional_fit <- mirt::mirt(Rotter.one.IRT, model = 1, 
                           itemtype = '2PL', 
                           SE = TRUE, 
                           Se.type = 'MHRM',
                           technical = list(NCYCLES = 10000))

```

```{r mod-itm-fit}
##Model Fit
unidim_fit_indices <- mirt::M2(unidimensional_fit)

##Item fit
unidimensional_itemfit <- mirt::itemfit(unidimensional_fit,
                                  fit_stats = c("S_X2", "G2","Zh", "infit"))

unidimensional_params <- mirt::coef(unidimensional_fit, IRTpars = TRUE, simplify = TRUE, rawug = FALSE) 

unidimensional_items <- data.frame(unidimensional_params$items)

irt_parameters <- cbind(unidimensional_itemfit[,1], unidimensional_items[,c(1:2)], unidimensional_itemfit[, c(11:14)])

irt_marameters_names <- c("Item", "Discrimination", "Difficulty", "Signed-Chisquare", "df", "RMSEA", "p")

colnames(irt_parameters) <- irt_marameters_names
```

```{r irt-tab}
papaja::apa_table(irt_parameters, caption = "IRT parameters")
```


```{r personfit}
##Person-fit
unidimensional.personfit <- mirt::personfit(unidimensional_fit) 

unidimensional.personfit_model_misfits <- subset(unidimensional.personfit, Zh < -2)

rownames(unidimensional.personfit_model_misfits)

nrow(unidimensional.personfit_model_misfits)

personfit <- ggplot(unidimensional.personfit, aes(x=Zh)) + geom_histogram(binwidth=.5,col=I("black"),fill="#00A08799")+geom_vline(xintercept = -2)+
  labs( y="Number of Participants", x = "Zh Statistics")


personfit
```

```{r irt-plots}
#IRT Curves
## Conventionals
item_info <- mirt::plot(unidimensional_fit, type= "infotrace")
test_info <- mirt::plot(unidimensional_fit, type= "info")
occ_info <- mirt::plot(unidimensional_fit, type= "trace")

item_info
test_info
occ_info
```


## Bonus: publication quality IRT plots

```{r ggplot2, include=F}
apatheme=theme_bw()+
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank(),
        #text=element_text(family = "Helvetica"),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        axis.title.x = element_text(size = 15),
        axis.title.y = element_text(size = 15),
        plot.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title=element_blank(),
        axis.line.x = element_line(color='black'),
        axis.line.y = element_line(color='black'),
        panel.border = element_rect(color = "black",
                                    fill = NA,
                                    size = 1))

```

```{r bonus-pub-plots}
tic <- tabledown::ggtestinfo(Rotter.one.IRT, unidimensional_fit)+
  geom_area()+aes(fill =  "#E64B3599")+theme(legend.position = "none")+
  scale_fill_identity()+xlab(expression(theta))+ylab(expression(I(theta)))+apatheme
tic

item_info <- tabledown::ggiteminfo(unidimensional_fit, 1,6)+
  geom_area()+aes(fill =  "#E64B3599")+theme(legend.position = "none")+
  scale_fill_identity()+xlab(expression(theta))+ylab(expression(I(theta)))+apatheme

item_info
```

# References
